---
title: "Homework 3"
author: Na Yun Cho
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
data("ny_noaa")

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.6,
  fig.height = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```


# Problem 2


### Part 1) Load, tidy, and wrangle the data
```{r}
accel_df = read.csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minute",
    names_prefix = "activity_", 
    values_to = "activity_count") %>%
  mutate(weekday_vs_weekend = recode(day, Monday = "weekday", Tuesday = "weekday", Wednesday = "weekday", 
                                     Thursday = "weekday", Friday = "weekday", Saturday = "weekend", Sunday = 
                                     "weekend")) %>%
  relocate(weekday_vs_weekend, .after = day) %>%
  mutate(
    minute = as.numeric(minute),
    day = as.factor(day), 
    day = fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
    weekday_vs_weekend = as.factor(weekday_vs_weekend))

accel_df %>% view
```

Comment: The resulting dataset shows the accelerometer data collected on a 63 year-old male with BMI 25, who was diagnosed with congestive heart failure and was admitted to the Advaned Cardiac Care Center of Columbia University Medical Center. More specifically, the data show the activity counts per each minute every day during the five weeks the data was collected from. The variables in the resulting dataset are `r names(accel_df)`. The variable 'minute' indicates which minute of the 24-hour day the activity occurred starting at midnight, and 'activity_count' denotes the count of the activity that occurred per each minute. In this dataset, there are a total of `r nrow(accel_df)` observations. 

### Part 2) Create a table that shows the total minutes of activity for each day 
```{r}
accel_df %>% 
  group_by(day , week) %>%
  summarize(act_total = sum(activity_count)) %>%
  pivot_wider(
    names_from = day, 
    values_from = act_total)
```

Comment: The resulting table shows the total minutes of activity per each day for five weeks. It seems that generally the days around weekends tend to show more peak activity counts than the weekdays. This is shown by activity counts over 600,000 occurring on week 1 of Sunday, week 3 of Monday, week 5 of Friday, and week 2 of Saturday whereas such counts never occurred on Tuesday, Wednesday, or Thursday. 
However, further analyses would have to be done to identify specific trends with certainty, because the activity counts seem to be quite similar across all days throughout each week.
Furthermore, the table shows that two Saturdays have significantly low numbers of activity counts compared to activity counts of all the other days. From inspection of the dataset, all the activity counts of these two Saturdays are recorded as '1', which could be due to mistake or errors. 

### Part 3) Create a single-panel plot that shows the 24-hour activity time courses for each day
```{r}
accel_df %>%
  ggplot(aes(x = minute, y = activity_count, group = day_id, color = day)) +
  geom_line(alpha = 0.2) + 
  geom_smooth(aes(group = day), se = FALSE)

```

Comment: The resulting graph shows the activity counts over 24-hours of each day during the five weeks. It seems that the activity counts generally cluster the most around 11:00am to 12pm (before and around 750 minutes) and around 8pm to 9:30 pm (before and around 1250 minutes). 
In addition, there seems to be a peak in activity on Sunday around 11am (before 750 minutes) and a peak in activity on Friday around 9pm (around 1250 minutes). 
Furthermore, according to the graph, I may be able to infer that the person usually wakes up before 8am (before 500 minutes) and goes to sleep around 10pm (around 1320 minutes). 


# Problem 3

### Part 1
```{r}
ny_noaa_df =
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax/10,
    tmin = tmin/10,
    prcp = prcp/10
  )

ny_noaa_df %>%
  count(snow) %>%
  arrange(desc(n))
```

Comment (before data cleaning): These data are accessed from the NOAA National Climatic Data Center, and the data show the amount of precipitation, snowfall, the depth of snow, and the maximum an minimum temperatures on days ranging from January 1st 1981 to December 31st 2010 within New York State. The variables included in this dataset are `r names(ny_noaa_df)`. The 'id' variable indicates the weather station ID, 'prcp' indicates the amount of precipitation in tenths of mm, 'snow' denotes snowfall in mm, 'snwd' denotes snow depth in mm, and 'tmax' and  'tmin' indicate the maximum and minimum temperatures in tenths of degrees Celsius respectively. The dataset has `r nrow(ny_noaa_df)` rows and `r ncol(ny_noaa_df)` columns. 
Some of the key variables of this dataset would be 'snow', 'tmax', and 'tmin', but I can observe that too many of the 'tmax', 'tmin', and 'snow' values are missing (written as NA's), which could affect the validity of the conclusions drawn from this dataset. 


Comment: The most commonly observed values for snowfall was 0, because the dataset contains 2008508 observations of the value 0 among all the values of snowfall. This may be due to the fact that it does not snow during most of the days in New York State. 


### Part 2
```{r}
ny_noaa_df %>% 
 filter(month %in% c("01", "07")) %>%
  group_by(id, year, month) %>%
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_tmax, group = id)) + 
  geom_point(alpha = 0.5) +
  geom_path() +
  facet_grid(.~month) +
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))
  
```

Comment: The two-panel plot shows the average maximum temperature in January and in July in each station throughout the years of 1981 to 2010. The plot shows a slight increasing trend in the average maximum temperature in January across the years, but it is hard to observe such a trend in the average maximum temperature in July across the years. 
Although the maximum temperatures are fairly consistent in January and July throughout the years, there seems to be more variation in maximum temperatures in January than in July across the years. 
In addition, the average maximum temperatures in January is much lower than the average maximum temperatures in July. 
For the data in January, there is one extreme outlier in 1982 that is lower than the values of the other years. 
For the data in July, there is also a drastic outlier that occurred in 1988 that is lower than the vaules of the other years. In both January and July, there are some other less drastic outlier as well. 


### Part 3
```{r}
tmax_vs_tmin = 
  ny_noaa_df %>%
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()

snowfall_dist = 
  ny_noaa_df %>% 
  filter(snow < 100) %>%
  filter(snow > 0) %>%
  ggplot(aes(x = year, y = snow)) + geom_violin(alpha = 0.5) + 
  theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))

tmax_vs_tmin + snowfall_dist
```

Comment: The two-panel plot includes the (i) hex plot shoing tmax vs tmin for the dataset and (ii) violin plot showing the distribution of snowfall values greater than 0 and less than 100 for every year. 

The hex plot indicates that although there is some variability throughout the dataset, most of the data cluster around the center of the distribution. There also seems to be some cases where the maximum temperature is lower than the minimum temperature, which would reduce the accuracy of the conclusions drawn from the dataset.

The violin plot shows that most of the annual snowfalls are between 0mm and 35mm across the years. The plot also indicates that annual snowfalls also tend to cluster around the values of 50 mm and 70 mm of snow across the years. These clusters are fairly consistent throughout the years. 